Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-20 VMCAI-revised-version/final.tex
effectivity

In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors. More precisely, we say that a WBTS $\mathscr{S}=(S, \rightarrow, \leq)$ is 
\begin{itemize}
%\item {\em effective} if there exists a pair of algorithms
%($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
%$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
%$M_\leq$ the ordering relation “$\leq$”.
%
\item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
exists an algorithm that computes $|\post(s)| \in \N \cup \{\infty\}$ 
on input $s \in S$.
%
%\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
%to the encoding of the ideal $\downarrow s$ is computable.
%
\item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning $\mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
%\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
%
\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.
%
%and there exist an algorithm $M_\downarrow$ that computes 
% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
% $s_1, s_2, \ldots, s_m \in S$.
%
\item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \post^*(s)$.
%
\item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
\end{itemize}

@@

As an example, let us remark that $(\N^d,\leq)$ is an ideally-effective wqo and reset-VASS are $\post$-effective, $\post$-ideal-effective and $\pred$-basis-effective but they are not $\post^*$-basis-effective (VASS are $\post^*$-basis-effective).

>>>>

In order to construct algorithms, we also require effective hypotheses about the computations of certain sets of predecessors and successors. More precisely, we say that an ordered transition system $\mathscr{S}=(S, \rightarrow, \leq)$ is 
\begin{enumerate}
%\item {\em effective} if there exists a pair of algorithms
%($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
%$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
%$M_\leq$ the ordering relation “$\leq$”.
%
\item {\em $\post$-effective} if $\mathscr{S}$ is effective, and if there
exists an algorithm that computes $|\post(s)| \in \N \cup \{\infty\}$ 
on input $s \in S$.
%
%\item {\em ideally-effective} \cite{BFM-ic17} if the function mapping the encoding of a state $s$
%to the encoding of the ideal $\downarrow s$ is computable.
%
\item {\em $\post$-ideal-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning $\mathop{\downarrow} \post(I)$, expressed as a finite union of (maximal) ideals.
%\item {\em completion-$\post$-effective} if its completion $\hat{\mathscr{S}}=(Ideals(S),\rightarrow, \subseteq)$ is $\post$-effective, i.e., that $\downarrow \post(I)$ is computable, expressed as a finite union of ideals, for all $I \in Ideals(S)$.
%
\item {\em $\pred$-basis-effective} \cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \pred(\mathop{\uparrow} s)$.
%
%and there exist an algorithm $M_\downarrow$ that computes 
% $\downarrow s$, on input $s \in S$, and some algorithm $M_{\mathop{\uparrow}^C}$ that computes the ideal decomposition of $S \setminus \mathop{\uparrow} \{ s_1, s_2, \ldots, s_m\}$, on input
% $s_1, s_2, \ldots, s_m \in S$.
%
\item {\em $\post^*$-basis-effective} \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any state $s \in S$ and returning a finite basis of $\mathop{\uparrow} \post^*(s)$.
%
\item {\em $\pred^*$-basis-effective} if $\mathscr{S}$ is effective and there exists an algorithm accepting
any ideal $I \in Ideals(S)$ and returning a finite basis of $\mathop{\downarrow} \pred^*(I)$.
\end{enumerate}

Counter machines are effective but don't enjoy any other effectivities among the list of five.
Reset VASS (hence VASS) and (front) lossy fifo automata are $\post$-effective, $\post$-ideal-effective, and $\pred$-basis-effective. There exist $\post$-effective WSTS that are not $\post$-ideal-effective (Proposition 35 in \cite{BFM-ic17}); there exist $\post$-ideal-effective WSTS that are not $\post$-effective (Proposition 36 in \cite{BFM-ic17}).There also exist WSTS that are not $\pred$-basis-effective (Proposition 45 in \cite{BFM-ic17}).
Reset VASS are not $\post^*$-basis-effective since $\post^*$-basis-effectiveness allows to compute the finite set of the set of minimal reachable states, hence it would allow to decide the zero-reachability problem that is undecidable for reset-VASS. However, VASS are $\post^*$-basis-effective \cite{DBLP:journals/corr/abs-2108-00889} and LCS (with $d$ fifo channels) are also $\post^*$-basis-effective because we have $(q,w_1,w_2,...,w_d) \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)  \in \mathop{\uparrow} \post^*(q_0,u_1,u_2,...,u_d)$ iff $(q,\epsilon,...,\epsilon)$ is reachable from $(q_0,u_1,u_2,...,u_d)$ that is decidable in LCS.
%
Reset-VASS are not $\pred^*$-basis-effective because computing a finite basis of $\mathop{\downarrow} \pred^*(q,0,...,0)$ would allow to decide whether $(q,0,...,0)$ is reachable that is undecidable for reset-VASS.

VASS are not $\pred^*$-basis-effective ????

@@

As an example, let us remark that $(\N^d,\leq)$ is an ideally-effective wqo.
